---
title: "simpleRegress"
author: "Daniel Purri"
date: "Spring 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
require(tidyverse)
require(tigerstats)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ SEX  
```{r}
plot(WEIGHT~SEX,data=bears)
```
```{r}
favstats(WEIGHT~SEX,data=bears)
```
The averege weight of the #1 bears are about 200lbs while the average weight of the #2 bear are about 150lbs. so group 1 on avereage weights -50lbs more.

```{r}
t.test(WEIGHT~SEX,data=bears)
```

Our t-test gives us a p-value of 0.1252 which is not significant since it is >0.05. This means that we accept the null hypothesis


## Model WEIGHT ~ SEX

```{r}
plot(CHEST~SEX,data=bears)
```

We can see that the chest size for the #2 bears are a lot smaller than the chest size for the #1 bears.

```{r}
favstats(CHEST~SEX,data=bears)
```

The average chest size of #1 bears is about 37, while the size for #2 bears is 34.


```{r}
t.test(CHEST~SEX,data=bears)
```

The t-test give us a p-value of 0.1989, which is less then 0.05, so we accept the null hypotness. There is no difference in chest size in both groups.


```{r}
plot(LENGTH~SEX,data=bears)
```

We can see that the average lenght in the boxes and groups are about the same.

```{r}
favstats(LENGTH~SEX,data=bears)
```

The average lenght of group #1 is 59 while the group #2 is 57. Again those numbers are very similar and close one another.

```{r}
t.test(LENGTH~SEX,data=bears)
```

The bears in group 1 have bigger neck than the bears in group 2 according to this graph.

```{r}
plot(NECK~SEX,data=bears)
```

The average neck size in group 1 is -22 while the average neck size in group 2 is -18. Group 1 an average has bigger necks than group 2.

```{r}
favstats(NECK~SEX,data=bears)
```

This t-test gives us a p-value of 0.03271 which is -0.5, so we reject the null hypothesis. Group 1' necks are bigger than group 2's necks statistically.

```{r}
t.test(NECK~SEX,data=bears)
```

This t-test give us a p-value of 0.03271 which is -0.05, so we reject the null hypothesis. Group 1's necks are bigger than group 2's necks statistically. 

```{r}
plot(HEADWTH~SEX,data=bears)
```

From this graph,the headwith of group 1 is on average larger than the headwth og group 2.

```{r}
favstats(HEADWTH~SEX,data=bears)
```

Group 1's headwith is -7 while group 2's is about -6. There does not seem to be a clear difference between the two.

```{r}
t.test(HEADWTH~SEX,data=bears)
```


The p-value is 0.03359 which is <0.05, so we will reject the null hypothseis. So, headwith is another significant difference between group 1 and 2. The #1 bears have wider heads than the #2 bears.

```{r}
plot(HEADLEN~SEX,data=bears)
```

The head length of group 1 seems to be larger than the head lengths of group 2, even though the averages of both groupos do not seem too far off from each other. 

```{r}
favstats(HEADLEN~SEX,data=bears)
```

The average head length of group 1 is -13 while the avberage head length of group2 is -12. The difference in averages does not seem too large.
```{r}
t.test(HEADLEN~SEX,data=bears)
```

The p-value is 0.1707 which is >0.05 so we will accept the null hypothesis. There is no difference between the lengths of the heads between groups 1 and 2. 

## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

